---
output:
  github_document:
    toc: true
    toc_depth: 4

---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sdmShiny

<!-- badges: start -->
<!-- badges: end -->

Last update:
```{r}
Sys.time()
```


Follow this link to go to the Shiny application:
https://anderskolstad.shinyapps.io/sdmShiny/


## Project homepage
All the documentation for this project is found on [this webpage](https://anders-kolstad.github.io/sdmShiny/).

## Installation
Developers should clone the repo and work from there.

## About the repo
Large files around 100mb or greater, or unessential raster files etc., are in the data/large/ folder whihc is not pushed (it's in the .gitignore file) and therefore only exists locally with Anders. The same for RData files intill I find a way to load them without getting magic number errors (readRDS don't work with knitr). R markdown files in the root folder produces html files that goes to the docs folder - these make up the github pages website (except the README file which makes up this page). 


