---
output:
  github_document:
    toc: true
    toc_depth: 4

---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sdmShiny

<!-- badges: start -->
<!-- badges: end -->

Last update:
```{r}
Sys.time()
```
Please see the [project homepage](https://anders-kolstad.github.io/sdmShiny/) for a full documentation and demonstration of the sdmShiny project.


## Installation
Developers should clone the repo and work from there.

## Example application
Follow this link
https://anderskolstad.shinyapps.io/demoSDM/


# Documentation
All the documentation for this project is found on [this webpage](https://anders-kolstad.github.io/sdmShiny/).

# About the repo
Large files around 100mb or greater, or unessential raster files etc., are in the data/large/ folder whihc is not pushed (it's in the .gitignore file) and therefore only exists locally with Anders. The same for RData files intill I find a way to load them without getting magic number errors (readRDS don't work with knitr). R markdown files in the root folder produces html files that do to the docs folder. These make up the github pages website (except the README file which makes up this page). 


