---
title: "sdmShiny-documentation"
author: "Anders L. Kolstad"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sdmShiny-documentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup}
#library(sdmShiny)
#source("/home/anders/Documents/R/Git projects/sdmShiny/R/spList.R")
```

# About
This vignette explains the workflow that ended up the the shiny app. The r project is arranged as an r package. All r code such as functions are in the 'data' folder.

# Get species list
This is the first function. It produces a list of species that we will later use to harvest occurence data from gbif. ...
```{r, eval = F}
mySpecies <- sl()
head(mySpecies)
# alternatively
# mySpecies(df = TRUE)
```

# Get environmental data (explanatory variables)
A .gri file with already collated environmental data was downloaded from NTNU box: https://ntnu.app.box.com/s/wcmr0dgoyz2yu6ielw6er1pm7h0gaisa/file/393633279036

```{r}
env <- raster::stack("data/large/PredictorVariables.grd")
names(env)
```

